---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
     
    library(dplyr)    
    
    library(ggplot2)

    activity <-  read.csv("activity.csv")
    clean_activity <- activity[!is.na(activity$steps),]
    
```
## What is mean total number of steps taken per day?
```{r}

    Total_steps <- aggregate(steps ~ date , data = clean_activity, FUN = sum)

    with(Total_steps, hist(steps, col = "Red", main = "Total number of steps taken each day"))
    
```

Mean and Median total number of steps taken per day
```{r}

    Steps_mean <- aggregate(steps ~ date , data = clean_activity, FUN = mean)
    Steps_median <- aggregate(steps ~ date , data = clean_activity, FUN = median) 
    
    step_day <- merge(Steps_mean, Steps_median, by = "date")
    colnames(step_day)[2:3] <- c("mean","median") 
    
    print(step_day[1:5,])

```



## What is the average daily activity pattern?

```{r}

    Steps_interval <- aggregate(steps ~ interval , data = clean_activity, FUN = mean)
    with(Steps_interval,  plot(interval, steps, type = "l"))

```

Intervals with the maximum number of steps

```{r}

    print(Steps_interval[order(-Steps_interval$steps)[1:3],])

```



## Imputing missing values

Total number of missing values

```{r}

    na_steps <- is.na(activity$steps)
    
    na_total <- sum(na_steps)
    
    print(na_total)
    
```
   
   Data after replacing the NA's values with the average of steps daily
    
```{r}
    na_index <- which(na_steps) ## Get's the index of the NA's

    
    num_cycles <- na_total / length(Steps_interval$steps)
    intial_Index <- 1
    
    for (counter in c(1:num_cycles)) {    
        
        subs_index <- na_index[intial_Index:(intial_Index+287)]    ## Extract the index to replace
        
        activity[subs_index,"steps"] = Steps_interval[1:288,"steps"]
        
        intial_Index <- intial_Index + 288    ## the next bloc to replace
    }
    
    Total_steps_NA <- aggregate(steps ~ date , data = activity, FUN = sum)
    
    par(mfrow= c(1,2))
    with(Total_steps_NA, hist(steps, col = "Red", main = "Data with the new values"))
    
    with(Total_steps, hist(steps, col = "Red", main = "Data with the old values"))
```


```{r}

    Steps_mean_NA <- aggregate(steps ~ date , data = activity, FUN = mean)
    Steps_median_NA <- aggregate(steps ~ date , data = activity, FUN = median) 
    
    par(mfrow= c(2,2))

    with(Steps_mean_NA ,hist(steps, main = "Mean with the new values") )
    with(step_day ,hist(mean, main = "Mean with the old values"))
    with(Steps_median_NA ,hist(steps, main = "Median with the new values"))
    with(step_day ,hist(median, main = "Median with the old values"))
    
    device.off()
    
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}

    activity$date <- as.Date(activity$date,"%Y-%m-%d")
    
    days <- weekdays(activity$date, abbreviate = "TRUE")
    
    weekends <- c("Sun","Sat")
    days[!(days %in% weekends)] <- "weekday"
    days[(days %in% weekends)] <- "weekend"
    
    activity$weekday <- as.factor(days)
    
 
    Steps_week <- aggregate(steps ~ weekday + interval, data = activity, FUN = mean)

    gplot <- ggplot(Steps_week, aes(interval, steps)) + facet_grid(weekday ~ .) + geom_line()

    print(gplot)
  
    
```


